*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-04 Code/Manuscript_O2.Rmd
edits

knitr::write_bib(c("knitr", "tidyverse", "lubridate", "ggplot2", "broom", "grid", "bookdown", "glue", "cowplot", "ggpubr", "ggh4x", "png", "Studio", "zoo", "photobiology", "photobiologyWavebands", "data.table", "googlesheets4",  "mgcv", "minpack.lm",  "tagger", 'RPackages.bib'))

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using *tidyverse* package [@R-tidyverse] running under R v4.1.3 and RStudio v2023.06.0 XXX Naaman to get location (https://posit.co/)  XXX. The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen *Prochlorococcus* strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

When assessing the presence of a particular protein complex at a sampling location, all peptides belonging to all subunits of the complex were included to give the greatest number of data points. As this data was acquired on a discovery mission rather than through targeted peptide approaches, it is difficult to discern accuracies of strain assignment annotations, particularly as the proteins of interest in this study are highly conserved across strains. We are, however, confident in clade classifications for each protein examined. A caveat to interpretation of this data is the peptide detection bias inherent to mass spectrometry CITATION. The data is also limited by the number and nature of protein spectra in the SEQUEST database: a peptide sequence was not determined unless there is already a known spectrum for that peptide in the SEQUEST database, hence some peptides of interest may not be identifiable. Furthermore, a peptide must be detected above a certain threshold abundance in order to be considered an accurate ‘hit’.

@@

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~) gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc., Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience, Germany) inserted into tubes for real-time measurements, with a temperature probe in
the bath of the bioreactor to correct [O~2~] measures for temperature fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05% CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

Figures \@ref(fig:OverlayPlots) provide visual representations of PUR, the black solid line and shaded area, in relation to the imposed PAR, the dotted line, under each imposed spectral wavebands for *P. marinus* MED4, SS120 and MIT9313, respectively. Figure (\@ref(fig:PurParPlot)) shows the relationship between calculated PUR versus imposed PAR for each *P. marinus* and each spectral waveband.

@@

 We filtered the dataset of Omar *et al*. [@omarAnnotationGenesEncoding2023], for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in *P. marinus* strains (MED4, MIT9313, SS120, and NATL2A). We grouped orthologs together by EC number and their KO number and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. Remove this once added XXXX
 XXXXADD METHODS for DNA REpair and for light dependent EnzymesXXXX

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from clade LLII/III (ex. SS120) and clade LLIV (ex. MIT9313) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. The exclusive assignment of ATP Synthase peptides to Clade IV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

XXXXCheck text for figureXXXX

@@

XXXRewrite caption if we change figure versionXXX
XXX Figure changed- need to rewrite caption if keeping this figure XXX
```{r CladeProchloroPhotosynthDepthO2, fig.cap = CladeProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus* clades: HLI (includes *P. marinus* MED4), HLII, LLI, LLII/III (includes P. marinus* SS120) and LLIV (includes *P. marinus* MIT9313).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum cumulative diel PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment (Figure \@ref(fig:PurParPlot)), despite being derived from the same photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. This bias in the range of data leads us to caution in comparing model fits of growth in response to cumulative diel PUR under red vs. blue wavebands.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)A). In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)A).  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)A), consistent with Murphy *et al*. [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct fits for different photoperiods refer to figure \@ref(fig:PhotoperiodPurFitsPlots)A.

@@

The representative of the LLII/III clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B). SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~.  Under 25 µM O~2~, red light again generated more growth of SS120 per unit cumulative diel PUR, than did blue light, again consistent with lower cost of growth through lower photoinactivation under red light (Figure \@ref(fig:BluevsRedPurFitsPlots)B). For distinct fits for different photoperiods refer to figure \@ref(fig:PhotoperiodPurFitsPlots)B.

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, lower, flatter response to cumulative diel PUR under 2.5 µM O~2~ (Figure \@ref(fig:PhotoperiodPurFitsPlots)C). Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands(Figure \@ref(fig:PhotoperiodPurFitsPlots)C).  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)C).

As with MED4 and SS120, our data again support enhanced growth under conditions of low cumulative diel PUR and 660 nm (red) spectral bandwidth, consistent with Murphy *et al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue, wavebands (Figure \@ref(fig:BluevsRedPurFitsPlots)C).  Interestingly, this protective effect of red light disappears for MIT9313 growing under 2.5 µM O~2~, possibly because photoinactivation is strongly suppressed under this low [O~2~]. For distinct fits for different photoperiod fits refer to figure \@ref(fig:PhotoperiodPurFitsPlots)C.

@@

 BluevsRedPurFitsPlots_cap <- glue("**chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, p value < 0.05) or dashed lines (not significantly different fits, p value > 0.05) using one-way ANOVA.",
                       .sep = " ")

@@

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* slr1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

FtsHTable <- tibble(Organism = c("*Prochlorococcus marinus*", "*Synechocystis sp. PCC6803*", "Function"),
                    FtsH1 = c("FtsH1", "SlrO228", "PSII Repair"),
                    FtsH2 = c("FtsH2", "Slr1604", "PSII Repair"),
                    FtsH3 = c("FtsH3", "Slr1463", "PSI biogenesis?"),
                    FtsH4 = c("FtsH4", "Slr1390", "Cell viability")
                    )

@@

\newpage
XXXXEdit text depending upon which ocean protein figures we useXXXXX
Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLII/III SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR  (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313 (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)).  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~] (Figure \@ref(fig:NATLProchloroProteinMetabDepthO2)), suggesting limited requirement for protein turnover under low [O~2~].

@@

readRDS(file.path(DataIn,"ProchlorococcusEnzymeTable.Rds")) %>% 
  knitr::kable(.,
             caption = "Enzyme Commission numbers (EC) and corresponsing enzyme names.",
            col.names =  c("EC",
                          "Enzyme Name",
                          "Kegg Orthology")
             )

>>>>

knitr::write_bib(c("knitr", "tidyverse", "lubridate", "ggplot2", "broom", "grid", "bookdown", "glue", "cowplot", "ggpubr", "ggh4x", "png", "zoo", "photobiology", "photobiologyWavebands", "data.table", "googlesheets4",  "mgcv", "minpack.lm"), 'RPackages.bib')

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using *tidyverse* package [@R-tidyverse] running under R v4.1.3 and RStudio v2023.06.0 [@RStudio]. The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen *Prochlorococcus* strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 [@bonisteelStrainSpecificDifferences2018].  Data for each strain was plotted against depth and  [O~2~] and sampling station.

When assessing the presence of a particular protein complex at a sampling location, all peptides belonging to all subunits of the complex were included to give the greatest number of data points. As this data was acquired on a discovery mission rather than through targeted peptide approaches, it is difficult to discern accuracies of strain assignment annotations, particularly as the proteins of interest in this study are highly conserved across strains. We are, however, confident in clade classifications for each protein examined. A caveat to interpretation of this data is the peptide detection bias inherent to mass spectrometry XXX Aurora CITATION XXX. The data is also limited by the number and nature of protein spectra in the SEQUEST database: a peptide sequence was not determined unless there is already a known spectrum for that peptide in the SEQUEST database, hence some peptides of interest may not be identifiable. Furthermore, a peptide must be detected above a certain threshold abundance in order to be considered an accurate ‘hit’.

@@

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~) gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc., Kingston, Ontario, Canada). [O~2~] *in situ* was verified using oxygen optodes (PyroScience, Germany) inserted into tubes for real-time measurements, with a temperature probe in
the bath of the bioreactor to correct [O~2~] measures for temperature fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05% CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

Figure \@ref(fig:OverlayPlots) provide visual representations of PUR, the black solid line and shaded area, in relation to the imposed PAR, the dotted line, under each imposed spectral wavebands for *P. marinus* MED4 (A,B,C), SS120 (D,E,F) and MIT9313 (G,H,I). Figure (\@ref(fig:PurParPlot)) shows the relationship between calculated PUR versus imposed PAR for each *P. marinus* and each spectral waveband.

@@

We filtered the dataset of Omar *et al*. [@omarAnnotationGenesEncoding2023], for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~; EC numbers identified by BRENDA as being activated, or inhibited by light; and EC numbers annotated by BioCyc [@karpBioCycCollectionMicrobial2019] as corresponding to the Gene Ontology Term (GO:0006281 - DNA repair), in *P. marinus* strains (MED4, MIT9313, SS120, and NATL2A). We grouped orthologs together by EC number and their KO number and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. Gene counts for Flavodiirons were obtained from Allahverdiyeva *et al*. [@allahverdiyevaCyanobacterialOxygenicPhotosynthesis2015a], as they do not have allocated EC numbers. A full list of enzymes and corresponding EC and KO numbers can be found in Table \@ref(tab:ProchlorococcusEnzymeTable).

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from clade LLII/III (ex. SS120) and clade LLIV (ex. MIT9313) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. The exclusive assignment of ATP Synthase peptides to clade LLIV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

@@

```{r CladeProchloroPhotosynthDepthO2, fig.cap = CladeProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus* clades: HLI (represents *P. marinus* MED4 annotated with solid black circles), LLI, LLII/III (represents *P. marinus* SS120 annotated with solid black circles) and LLIV (represents *P. marinus* MIT9313 annotated with solid black circles).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum cumulative diel PUR under blue light is almost 3 times that of white LED light, and about 5 times the red light treatment (Figure \@ref(fig:PurParPlot)), despite being derived from the same photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. This bias in the range of data leads us to caution in comparing model fits of growth in response to cumulative diel PUR under red vs. blue wavebands.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)A,B,C). In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A,B), and no evidence of inhibition of growth at any achieved cumulative diel PUR.  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)A,B), consistent with Murphy *et al*. [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct fits for different photoperiods refer to figure \@ref(fig:PhotoperiodPurFitsPlots)A,B and C.

@@

The representative of the LLII/III clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)F). Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:PhotoperiodPurFitsPlots)F). SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:PhotoperiodPurFitsPlots)F).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~.  Under 25 µM O~2~, red light again generated more growth of SS120 per unit cumulative diel PUR, than did blue light, again consistent with lower cost of growth through lower photoinactivation under red light (Figure \@ref(fig:BluevsRedPurFitsPlots)E). For distinct fits for different photoperiods refer to figure \@ref(fig:PhotoperiodPurFitsPlots)D,E and F.

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, lower, flatter response to cumulative diel PUR under 2.5 µM O~2~ (Figure \@ref(fig:PhotoperiodPurFitsPlots)G,H). Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands (Figure \@ref(fig:PhotoperiodPurFitsPlots)G).  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ (Figure \@ref(fig:PhotoperiodPurFitsPlots)H).  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)I).

As with MED4 and SS120, our data again support enhanced growth under conditions of low cumulative diel PUR and 660 nm (red) spectral bandwidth, consistent with Murphy *et al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue, wavebands (Figure \@ref(fig:BluevsRedPurFitsPlots)G,H).  Interestingly, this protective effect of red light disappears for MIT9313 growing under 2.5 µM O~2~, possibly because photoinactivation is strongly suppressed under this low [O~2~] (Figure \@ref(fig:BluevsRedPurFitsPlots)I). For distinct fits for different photoperiod fits refer to figure \@ref(fig:PhotoperiodPurFitsPlots)G,H and I.

@@

 BluevsRedPurFitsPlots_cap <- glue("**chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4 (A,B,C), SS120 (D,E,F) and MIT9313 (G,H,I).",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, p value < 0.05) or dashed lines (not significantly different fits, p value > 0.05) using one-way ANOVA.",
                       .sep = " ")

@@

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* *slr*1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

FtsHTable <- tibble(Organism = c("*Prochlorococcus marinus*", "*Synechocystis sp. PCC6803*", "Function"),
                    FtsH1 = c("FtsH1", "*Slr*0228", "PSII Repair"),
                    FtsH2 = c("FtsH2", "*Slr*1604", "PSII Repair"),
                    FtsH3 = c("FtsH3", "*Sll*1463", "PSI biogenesis?"),
                    FtsH4 = c("FtsH4", "*Slr*1390", "Cell viability")
                    )

@@

@@

readRDS(file.path(DataIn,"ProchlorococcusEnzymeTable.Rds")) %>% 
  select(EnzymeName, EC, KO) %>% 
  knitr::kable(.,
             caption = "Enzymes shown in Figures \\@ref(fig:ProchlorococcusEnzymeKms),  \\@ref(fig:DNARepairFig) and \\@ref(fig:ProchlorococcusLightEnzymes) with their Enzyme Commission numbers (EC) and Kegg Orthology (KO).",
            col.names =  c(
                          "Enzyme Name",
                          "EC",
                          "Kegg Orthology")
             )
